---
title: "State Service Los Frequency of Use Table"
author: "Connor Flynn"
date: "12/29/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(haven)
library(feather)
library(arrow)
library(chorddiag)
library(circlize)
library(kableExtra)
library(urbnmapr)
library(forcats)
library(sf)
options(scipen = 99)
```

Load in Data

14 million treatment episodes from 2012-2020

```{r}
teds_d <- arrow::read_feather(here("data/tedsD_2012_2020.feather"))
```


Clean Column Names

```{r}
teds_d <- teds_d %>% 
  clean_names()
```


Filter for years with admission/discharge data (2015-2020, 9 million records)


```{r}
teds_d_select_years <- teds_d %>% 
  filter(disyr %in% c("2015", "2016", "2017", "2018", "2019", "2020"))
```

Count records per year

```{r}
table(teds_d_select_years$disyr)
```

*Filter for Completed Treatments ONLY* (4 million treatment episodes)

```{r}
teds_d_select_years_completed <- teds_d_select_years %>% 
  filter(reason == "1")
```


Replace -9 with NA

```{r}
teds_d_select_years_completed[teds_d_select_years_completed == "-9"] <- NA
```


Relevant variables to decrease size of large dataframe

```{r}
teds_d_select_years_completed_relevant <- teds_d_select_years_completed %>% 
  select(stfips, freq1_d, services_d, los)
```


Bin LOS
- easier for data visualization
- binned to limit number of classes and not bias random forest

```{r}
teds_d_select_years_completed_relevant$los_binned <- cut(
  teds_d_select_years_completed_relevant$los,
  breaks = c(0, 1, 5, 10, 20, 30, 31, 32, 33, 34, 35, 36, 37),
  labels = c('1 day', '2-5 days', '6-10 days', '11-20 days', '21-30 days', '31-45 days', '46-60 days', '61-90 days', '91-120 days', '121-180 days', '181-365', '>365 days'),
  include.lowest = TRUE
)

# Check the frequency table of the new variable
teds_d_select_years_completed_relevant %>% 
  select(los, los_binned)
```
```{r}
#teds_d_select_years_completed_relevant[is.na(teds_d_select_years_completed_relevant)] <- "unknown"
```

Replace frequency of use / services codes with meaning

```{r}
teds_d_select_years_completed_relevant$freq1_d[teds_d_select_years_completed_relevant$freq1_d == "1"] <- "no use"


teds_d_select_years_completed_relevant$freq1_d[teds_d_select_years_completed_relevant$freq1_d == "2"] <- "some use"


teds_d_select_years_completed_relevant$freq1_d[teds_d_select_years_completed_relevant$freq1_d == "3"] <- "daily use"


teds_d_select_years_completed_relevant$services_d[teds_d_select_years_completed_relevant$services_d == "1"] <- "Detox, 24-hour, hospital inpatient"

teds_d_select_years_completed_relevant$services_d[teds_d_select_years_completed_relevant$services_d == "2"] <- "Detox, 24-hour, free-standing residential"

teds_d_select_years_completed_relevant$services_d[teds_d_select_years_completed_relevant$services_d == "3"] <- "Rehab/residential, hospital (non-detox)"

teds_d_select_years_completed_relevant$services_d[teds_d_select_years_completed_relevant$services_d == "4"] <- "Rehab/residential, short term (30 days or fewer)"

teds_d_select_years_completed_relevant$services_d[teds_d_select_years_completed_relevant$services_d == "5"] <- "Rehab/residential, long term (more than 30 days)"

teds_d_select_years_completed_relevant$services_d[teds_d_select_years_completed_relevant$services_d == "6"] <- "Ambulatory, intensive outpatient"

teds_d_select_years_completed_relevant$services_d[teds_d_select_years_completed_relevant$services_d == "7"] <- "Ambulatory, non-intensive outpatient"

teds_d_select_years_completed_relevant$services_d[teds_d_select_years_completed_relevant$services_d == "8"] <- "Ambulatory, detoxification"


```



Frequency of Use 

```{r}
frequency_of_use_by_state <- teds_d_select_years_completed_relevant %>% 
  group_by(stfips, freq1_d) %>% 
  count(name = "sum_freq_use_in_state")
```


Mutate Percent column 
- each state will have four rows for freq1_d that should sum to 100%
```{r}
frequency_of_use_by_state <- frequency_of_use_by_state %>% 
  group_by(stfips) %>% 
  mutate(percent_freq_use = sum_freq_use_in_state / sum(sum_freq_use_in_state) * 100)
```

Drop sum since it was only used to calculate the percentage
- change state and frequency of use to factors

```{r}
frequency_of_use_by_state <- frequency_of_use_by_state %>% 
  select(-sum_freq_use_in_state) 


frequency_of_use_by_state$stfips <- as.factor(frequency_of_use_by_state$stfips)
frequency_of_use_by_state$freq1_d <- as.factor(frequency_of_use_by_state$freq1_d)
```


```{r}
# frequency_of_use_by_state_wider <- frequency_of_use_by_state %>% 
#   pivot_wider(names_from = freq1_d, values_from = percent_freq_use)
```



Services by State

```{r}
services_by_state <- teds_d_select_years_completed_relevant %>% 
  group_by(stfips, services_d) %>% 
  count(name = "sum_service_in_state")
```

Mutate Percent column 
- each state will have some number of rows for each service (whichever are offered in state) that should sum to 100%

```{r}
services_by_state <- services_by_state %>% 
  group_by(stfips) %>% 
  mutate(percent_in_service = sum_service_in_state / sum(sum_service_in_state) * 100)
```

Drop sum since it was only used to calculate the percentage
- change state and services at discharge to factors

```{r}
services_by_state <- services_by_state %>% 
  select(-sum_service_in_state) 


services_by_state$stfips <- as.factor(services_by_state$stfips)
services_by_state$services_d <- as.factor(services_by_state$services_d)
```

```{r}
# services_by_state_wider <- services_by_state %>% 
#   pivot_wider(names_from = services_d, values_from = percent_in_service)
```


LOS by State 

```{r}
los_by_state <- teds_d_select_years_completed_relevant %>% 
  group_by(stfips, los_binned) %>% 
  count(name = "sum_los_binned_in_state")
```

Mutate Percent column 
- each state will have 12 rows for each los bin  that should sum to 100%

```{r}
los_by_state <- los_by_state %>% 
  group_by(stfips) %>% 
  mutate(percent_in_los_bin = sum_los_binned_in_state / sum(sum_los_binned_in_state) * 100)
```

Drop sum since it was only used to calculate the percentage
- change state and los bin to factors

```{r}
los_by_state <- los_by_state %>% 
  select(-sum_los_binned_in_state)

los_by_state$stfips <- as.factor(los_by_state$stfips)
los_by_state$los_binned <- as.factor(los_by_state$los_binned)
```

```{r}
# los_by_state_wider <- los_by_state %>% 
#   pivot_wider(names_from = los_binned, values_from = percent_in_los_bin)
```


```{r}
# head(los_by_state_wider)
# head(services_by_state_wider)
# head(frequency_of_use_by_state_wider)
```

```{r}
head(los_by_state)
head(services_by_state)
head(frequency_of_use_by_state)
```

Plot to show concept for final tile map

```{r}
ggplot(frequency_of_use_by_state, aes(x = freq1_d, y = stfips, fill = percent_freq_use)) +
  geom_tile() +
  scale_fill_viridis_c() +
  theme_minimal()
```


Join Data


```{r}

# Join the dataframes
df_merged <- los_by_state %>%
  full_join(services_by_state, by = "stfips") %>%
  full_join(frequency_of_use_by_state, by = "stfips")


```

```{r}
df_merged_longer <- df_merged %>% 
  pivot_longer(cols = c("los_binned", "freq1_d", "services_d"),
               names_to = "importance_feature",
               values_to = "desription")

df_merged_longer2 <- df_merged_longer %>%
  mutate(percentage = case_when(
    importance_feature == "los_binned" ~ percent_in_los_bin,
    importance_feature == "freq1_d" ~ percent_freq_use,
    importance_feature == "services_d" ~ percent_in_service,
    TRUE ~ NA_real_
  )) 


df_merged_longer2 <- df_merged_longer2 %>% 
  select(stfips, importance_feature, desription, percentage) %>% 
  distinct()



# df_merged_longer <- df_merged_longer %>% 
#   pivot_longer(cols = c("percent_in_los_bin", "percent_freq_use", "percent_in_service"),
#                #names_to = "importance_feature",
#                values_to = "percentage")
```


```{r}
head(df_merged_longer2)
```
```{r}
df_merged_longer2$percentage <- round(df_merged_longer2$percentage, digit = 2)
```

Add state names based on fips codes

```{r}
state_info <- data.frame(
  stfips = c(1, 2, 4, 5, 6, 8, 9, 10, 11, 12, 13, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 42, 44, 45, 46, 47, 48, 49, 50, 51, 55, 56, 72),
  State = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Wisconsin", "Wyoming", "Puerto Rico")
)

df_merged_longer2 <- merge(df_merged_longer2, state_info, by = "stfips", all.x = TRUE)
```


Final tile map 

Check plot to make sure it matches after joining
1) los_by_state
2) services_by state
3) frequency_of_use_by_state

```{r}
tile_map <- ggplot(df_merged_longer2, aes(x = desription, y = State, fill = percentage)) +
  geom_tile() +
  facet_grid(~importance_feature, scales = "free_x") +
  theme_minimal() +
  scale_fill_viridis_c() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 3),
        axis.text.y = element_text(size = 3))+
  geom_text(aes(label = percentage), color = "white", size = 1) +
  labs(x = "",
       y = "",
       fill = "% of Treatment Episodes") 

tile_map
```

Make interactive

```{r}
ggplotly(tile_map)
```




Read in Partial Dependence Plot Data from TACC


```{r}
pdp_stfips <- read_csv(here("data/pdp_stfips.csv"))
```

Get state names 

```{r}
pdp_stfips_names <- merge(pdp_stfips, state_info, by = "stfips", all.x = TRUE)
```

Plot PDP

```{r}
ggplot(pdp_stfips_names, aes(y = reorder(State, yhat), x = yhat)) +
  geom_bar(stat = "identity", fill = "#639CA4FF", color = "black") +
  theme_minimal() +
  labs(title = "Partial Dependence Plot",
       x = "Prediction of No-Use at Discharge",
       y = "")
```

Interrogation

Nevada


```{r}
nevada <- teds_d_select_years_completed_relevant %>% 
  filter(stfips == "32")
```


```{r}
nevada_grouped <- nevada %>% 
  group_by(services_d, freq1_d) %>% 
  count()
```

```{r}
nevada_grouped <- nevada_grouped %>% 
  group_by(services_d) %>% 
  mutate(percent = n / sum(n) * 100)
```

Mass

```{r}
mass <- teds_d_select_years_completed_relevant %>% 
  filter(stfips == "25")
```


```{r}
mass_grouped <- mass %>% 
  group_by(services_d, freq1_d) %>% 
  count()
```

```{r}
mass_grouped <- mass_grouped %>% 
  group_by(services_d) %>% 
  mutate(percent = n / sum(n) * 100)
```



Louisianna

```{r}
louisianna <- teds_d_select_years_completed_relevant %>% 
  filter(stfips == "22")
```


```{r}
louisianna_grouped <- louisianna %>% 
  group_by(services_d, freq1_d) %>% 
  count()
```

```{r}
louisianna_grouped <- louisianna_grouped %>% 
  group_by(services_d) %>% 
  mutate(percent = n / sum(n) * 100)
```
