---
title: "Mapping Frequency of Use Change Binary"
author: "Connor Flynn"
date: "12/17/2023"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(haven)
library(feather)
library(arrow)
library(chorddiag)
library(circlize)
library(kableExtra)
library(urbnmapr)
library(forcats)
library(sf)
options(scipen = 99)

```

Load in Data

```{r}
teds_d <- arrow::read_feather(here("data/tedsD_2012_2020.feather"))
```


Clean Column Names

```{r}
teds_d <- teds_d %>% 
  clean_names()
```


Filter for years with admission/discharge data


```{r}
teds_d_select_years <- teds_d %>% 
  filter(disyr %in% c("2015", "2016", "2017", "2018", "2019", "2020"))
```



Relevant variables to decrease size of large dataframe

```{r}
teds_d_select_years_relevant <- teds_d_select_years %>% 
  select(reason, freq1, freq1_d, los, employ, employ_d, livarag, livarag_d, freq_atnd_self_help, freq_atnd_self_help_d, stfips, services_d, los)
```

Replace -9 with NA

```{r}
teds_d_select_years_relevant[teds_d_select_years_relevant == "-9"] <- NA
```



<!-- Mutate as character to replace na with unknown in next chunk -->

<!-- ```{r} -->
<!-- teds_d_select_years_relevant <- teds_d_select_years_relevant %>%  -->
<!--   mutate_all(as.character) -->
<!-- ``` -->


# ```{r}
# teds_d_select_years_relevant[is.na(teds_d_select_years_relevant)] <- "unknown"
# ```

Replace frequency of use / services codes with meaning

```{r}
teds_d_select_years_relevant$freq1_d <- as.character(teds_d_select_years_relevant$freq1_d)

teds_d_select_years_relevant$freq1_d[teds_d_select_years_relevant$freq1_d == "1"] <- "no use"


teds_d_select_years_relevant$freq1_d[teds_d_select_years_relevant$freq1_d == "2"] <- "some use"


teds_d_select_years_relevant$freq1_d[teds_d_select_years_relevant$freq1_d == "3"] <- "daily use"

teds_d_select_years_relevant$freq1_d[is.na(teds_d_select_years_relevant$freq1_d)] <- "unknown"

```


```{r}

teds_d_select_years_relevant$services_d <- as.character(teds_d_select_years_relevant$services_d)


teds_d_select_years_relevant$services_d[teds_d_select_years_relevant$services_d == "1"] <- "Detox, 24-hour, hospital inpatient"

teds_d_select_years_relevant$services_d[teds_d_select_years_relevant$services_d == "2"] <- "Detox, 24-hour, free-standing residential"

teds_d_select_years_relevant$services_d[teds_d_select_years_relevant$services_d == "3"] <- "Rehab/residential, hospital (non-detox)"

teds_d_select_years_relevant$services_d[teds_d_select_years_relevant$services_d == "4"] <- "Rehab/residential, short term (30 days or fewer)"

teds_d_select_years_relevant$services_d[teds_d_select_years_relevant$services_d == "5"] <- "Rehab/residential, long term (more than 30 days)"

teds_d_select_years_relevant$services_d[teds_d_select_years_relevant$services_d == "6"] <- "Ambulatory, intensive outpatient"

teds_d_select_years_relevant$services_d[teds_d_select_years_relevant$services_d == "7"] <- "Ambulatory, non-intensive outpatient"

teds_d_select_years_relevant$services_d[teds_d_select_years_relevant$services_d == "8"] <- "Ambulatory, detoxification"

teds_d_select_years_relevant$services_d[is.na(teds_d_select_years_relevant$services_d)] <- "unknown"


```

Mutate all as factor

```{r}
# teds_d_select_years_relevant <- teds_d_select_years_relevant %>% 
#   mutate_all(as.factor)
```


*Select for only completed treatments*

```{r}
teds_d_select_years_relevant_completed <- teds_d_select_years_relevant %>% 
  filter(reason == "1")
```

Count the freq1_d classes for all states

```{r}
freq_use_states <- teds_d_select_years_relevant_completed %>% 
  group_by(stfips, freq1_d) %>% 
  count(name = "count") 
```

Calculate percentage (each state should equal 100)

```{r}
freq_use_states <- freq_use_states %>% 
  group_by(stfips) %>% 
  mutate(percentage = count / sum(count) * 100)
```

Bring in map data

```{r, include=FALSE}
#data = urbnmapr::states
```

Change fips to match map 

```{r}
#freq_use_states$stfips <- as.numeric(freq_use_states$stfips)
```


```{r, include=FALSE}
freq_use_states$stfips_recode <- sprintf('%02d', freq_use_states$stfips)
```

```{r}
freq_use_states %>%  
  select(stfips, stfips_recode) %>% 
  unique()
```



Change column name to match map

```{r, include=FALSE}
colnames(freq_use_states)[colnames(freq_use_states) == "stfips_recode"] <- "state_fips"
```

```{r}
#freq_use_states$state_fips <- as.character(freq_use_states$state_fips)
```


```{r}
states_map <- get_urbn_map(map = "states", sf = TRUE)
```


Join frequency ouf use - state data with mapping data
```{r, include=FALSE}


freq_use_states <- left_join(freq_use_states,
                          states_map,
                          by = "state_fips")
```

Filter for no use for zoomed in map


```{r}
freq_use_states_no_use <- freq_use_states %>% 
  filter(freq1_d == "no use")
```


```{r}
# breaks <- seq(0, 1, by = 0.1)
# labels <- c("0-10%", "11-20%", "21-30%", "31-40%", "41-50%", "51-60%", "61-70%", "71-80%", "81-90%", "91-100%")
```


Plot 


```{r}
ggplot(freq_use_states_no_use) +
  geom_sf(freq_use_states_no_use,
          mapping = aes(geometry = geometry, fill = percentage),
          color = "#ffffff", size = 0.25) +
  labs(fill = "% of Treatment Episodes") +
   coord_sf(datum = NA)+
  #geom_sf_text(data = get_urbn_labels(map = "territories_states", sf = TRUE), 
                #aes(label = state_abbv), 
            #size = 3) +
   theme_minimal() +
  theme(panel.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 4),  # Adjust the legend text size
    legend.title = element_text(size = 5),
    strip.text = element_text(size = 4))
```

Add color bins

```{r}

freq_use_states_no_use <- freq_use_states_no_use %>% 
  mutate(percentage_bin = cut(percentage, breaks=c(0, 10,20,30,40,50, 60, 70, 80, 90, 100)))

ggplot(freq_use_states_no_use) +
  geom_sf(mapping = aes(geometry = geometry, fill = percentage_bin),
          color = "#ffffff", size = 0.25) +
  labs(fill = "% of Treatment Episodes",
      title = "Completed Treatment Episodes with `no use` at Discharge",
      subtitle = "TEDS-D 2015-2020") +
  scale_fill_viridis_d(option = "D") +
  coord_sf(datum = NA) +
  theme_minimal() +
  theme(
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    legend.text = element_text(size = 4),
    legend.title = element_text(size = 5),
    strip.text = element_text(size = 4)
  ) 
  




```


```{r}
# Dynamically set breaks based on quantiles

# freq_use_states <- freq_use_states %>%
#   drop_na()


```



facet into 4 frequency of use classes


```{r}
ggplot(freq_use_states) +
  geom_sf(mapping = aes(geometry = geometry, fill = percentage),
          color = "#ffffff", size = 0.25) +
  labs(fill = "% of Treatment Episodes") +
   coord_sf(datum = NA)+
  #geom_sf_text(data = get_urbn_labels(map = "territories_states", sf = TRUE), 
                #aes(label = state_abbv), 
            #size = 3) +
   theme_minimal() +
  theme(panel.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 4),  # Adjust the legend text size
    legend.title = element_text(size = 5),
    strip.text = element_text(size = 4)) +
  scale_fill_viridis_c(option = "D") +
  facet_wrap(~freq1_d)
```

Add color scale


```{r}
freq_use_states <- freq_use_states %>% 
  mutate(percentage_bin = cut(percentage, breaks=c(0, 10,20,30,40,50, 60, 70, 80, 90, 100)))
```

```{r}
ggplot(freq_use_states) +
  geom_sf(freq_use_states,
          mapping = aes(geometry = geometry, fill = percentage_bin),
          color = "#ffffff", size = 0.25) +
  labs(fill = "% of Treatment Episodes",
       title = "Completed Treatment Episodes and Frequency of Use at Discharge",
       subtitle = "TEDS-D 2015-2020") +
   coord_sf(datum = NA)+
  #geom_sf_text(data = get_urbn_labels(map = "territories_states", sf = TRUE), 
                #aes(label = state_abbv), 
            #size = 3) +
   theme_minimal() +
  theme(panel.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 4),  # Adjust the legend text size
    legend.title = element_text(size = 5),
    strip.text = element_text(size = 4)) +
  scale_fill_viridis_d(option = "D") +
  facet_wrap(~freq1_d)
```


Staccked


```{r}

# Reorder the levels of freq1_d column
freq_use_states$freq1_d <- factor(freq_use_states$freq1_d, 
                                  levels = c("no use", "some use", "daily use", "unknown"))

# Remove duplicates in state_name based on the percentage where freq1_d == "no use"
distinct_states <- freq_use_states %>%
  filter(freq1_d == "no use") %>%
  arrange(percentage) %>%
  select(state_name) %>%
  distinct()

# Reorder state_name based on the percentage where freq1_d == "no use"
freq_use_states <- freq_use_states %>%
  arrange(freq1_d, percentage) %>%
  mutate(state_name = factor(state_name, levels = distinct_states$state_name))

# Plot with reversed fill order
ggplot(freq_use_states, aes(x = percentage, y = state_name, fill = forcats::fct_rev(freq1_d))) +
  geom_bar(position = "stack", stat = "identity") +
  theme_minimal() +
  scale_fill_viridis_d() +
  labs(fill = "Frequency of Use",
       y = "", 
       x = "Treatment Episodes (%)",
  title = "Frequency of Use at Discharge",
  subtitle = "TEDS-D Completed Treatment Episodes") +
  guides(fill = guide_legend(reverse = TRUE))
```



Calculate the percentage of treatment episodesthat fall into each service for each state

```{r}
services_by_state <- teds_d_select_years_relevant_completed %>% 
  group_by(stfips, services_d) %>% 
  count(name = "number_in_service")
  
```


```{r}
services_by_state <- services_by_state %>% 
  group_by(stfips) %>% 
  mutate(percent_in_service = number_in_service / sum(number_in_service) * 100)
```

```{r}
services_by_state <- services_by_state %>% 
  mutate(services_percentage_bin = cut(percent_in_service, breaks=c(0, 10,20,30,40,50, 60, 70, 80, 90, 100)))
```


```{r, include=FALSE}
services_by_state$stfips_recode <- sprintf('%02d', services_by_state$stfips)
```


```{r}
services_by_state %>% 
  select(stfips, stfips_recode)
```


```{r, include=FALSE}
colnames(services_by_state)[colnames(services_by_state) == "stfips_recode"] <- "state_fips"
```


```{r, include=FALSE}
services_by_state <- left_join(get_urbn_map(map = "states", sf = TRUE),
                          services_by_state,
                          by = "state_fips")
```

```{r}
# services_by_state <- services_by_state %>% 
#   drop_na()
```


```{r}
ggplot(services_by_state) +
  geom_sf(services_by_state,
          mapping = aes(geometry = geometry, fill = services_percentage_bin),
          color = "#ffffff", size = 0.25) +
  labs(fill = "% of Treatment Episodes",
       title = "Percentage of Completed Treatment Episodes in Services by State",
       subtitle = "TEDS-D 2015-2020") +
   coord_sf(datum = NA)+
  #geom_sf_text(data = get_urbn_labels(map = "territories_states", sf = TRUE), 
                #aes(label = state_abbv), 
            #size = 3) +
   theme_minimal() +
  theme(panel.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 4),  # Adjust the legend text size
    legend.title = element_text(size = 5),
    strip.text = element_text(size = 4)) +
  scale_fill_viridis_d(option = "D") +
  facet_wrap(~services_d)
```
Stacked

```{r}
ggplot(services_by_state, aes(x = percent_in_service, y = state_name, fill = forcats::fct_rev(services_d))) +
  geom_bar(position = "stack", stat = "identity") +
  theme_minimal() +
  scale_fill_viridis_d() +
  labs(fill = "Frequency of Use",
       y = "", 
       x = "Treatment Episodes (%)",
  title = "Frequency of Use at Discharge",
  subtitle = "TEDS-D Completed Treatment Episodes") +
  guides(fill = guide_legend(reverse = TRUE))
```





Median LOS by State

Colored Boxplot by %no use

```{r}
teds_d_select_years_relevant_completed$stfips_recode <- sprintf('%02d',teds_d_select_years_relevant_completed$stfips)

colnames(teds_d_select_years_relevant_completed)[colnames(teds_d_select_years_relevant_completed) == "stfips_recode"] <- "state_fips"

#teds_d_select_years_relevant_completed$state_fips <- as.numeric(teds_d_select_years_relevant_completed$state_fips)

teds_d_select_years_relevant_completed %>% 
  select(stfips, state_fips) %>% 
  unique()

```


```{r}
freq_use_states_no_geo <- freq_use_states %>% 
  st_drop_geometry()
```

Check on lost data

```{r}
teds_d_select_years_relevant_completed_freq_use_states <- right_join(teds_d_select_years_relevant_completed, freq_use_states_no_geo,
                                                                    by = c("state_fips", "freq1_d"))
```

```{r}
teds_d_select_years_relevant_completed_freq_use_states$los <- as.numeric(teds_d_select_years_relevant_completed_freq_use_states$los)
```


```{r}
teds_d_select_years_relevant_completed_freq_use_states <- teds_d_select_years_relevant_completed_freq_use_states %>% 
  group_by(state_name) %>% 
  mutate(median = median(los, na.rm = TRUE))
```

```{r}
ggplot(teds_d_select_years_relevant_completed_freq_use_states, aes(x = los, y = reorder(state_name, median))) +
  geom_boxplot(outlier.size = 0.1) +
  theme_minimal() +
  #scale_fill_viridis_c(option = "C") +
  theme(text = element_text(size=5),
        legend.position = "none") +
  labs(title = "Length of Stay by State",
       x = "Length of Stay (Days)",
       y = "State")
```

```{r}
ggplot(teds_d_select_years_relevant_completed_freq_use_states, aes(x = los, y = reorder(state_name, median))) +
  geom_boxplot(outlier.size = 0.1) +
  theme_minimal() +
  #scale_fill_viridis_c(option = "C") +
  theme(text = element_text(size=5),
        legend.position = "none") +
  labs(title = "Length of Stay by State",
       x = "Length of Stay (Days)",
       y = "State") +
  facet_grid(~freq1_d)
```

Check 

```{r}
teds_d_select_years_relevant_completed$los <- as.numeric(teds_d_select_years_relevant_completed$los) 

median_los_completed <- teds_d_select_years_relevant_completed %>% 
  group_by(stfips) %>% 
  mutate(median = median(los, na.rm = TRUE)) 
  
mean_los_completed <- teds_d_select_years_relevant_completed %>% 
  group_by(stfips) %>% 
  mutate(mean = mean(los, na.rm = TRUE)) 

teds_d_select_years_relevant$los <- as.numeric(teds_d_select_years_relevant$los) 

median_los<- teds_d_select_years_relevant %>% 
  group_by(stfips) %>% 
  mutate(median = median(los, na.rm = TRUE)) 
```



LOS Binned


```{r}
# Create a new variable 'LOS_binned'
teds_d_select_years_relevant_completed_freq_use_states$los_binned <- cut(
  teds_d_select_years_relevant_completed_freq_use_states$los,
  breaks = c(0, 1, 5, 10, 20, 30, 31, 32, 33, 34, 35, 36, 37),
  labels = c('1 day', '2-5 days', '6-10 days', '11-20 days', '21-30 days', '31-45 days', '46-60 days', '61-90 days', '91-120 days', '121-180 days', '181-365', '>365 days'),
  include.lowest = TRUE
)

# Check the frequency table of the new variable
teds_d_select_years_relevant_completed_freq_use_states %>% 
  select(los, los_binned)
```


```{r}
los_binned_by_state <- teds_d_select_years_relevant_completed_freq_use_states %>% 
  group_by(state_name, los_binned) %>% 
  count()
```

```{r}
los_binned_by_state <- los_binned_by_state %>% 
  group_by(state_name) %>% 
  mutate(percent_in_los = n / sum(n) * 100)
```

```{r}
ggplot(los_binned_by_state, aes(x = percent_in_los, y = state_name, fill = forcats::fct_rev(los_binned))) +
  geom_bar(position = "stack", stat = "identity") +
  theme_minimal() +
  scale_fill_viridis_d() +
  labs(fill = "Length of Stay",
       y = "", 
       x = "Treatment Episodes (%)",
  title = "LOS by State",
  subtitle = "TEDS-D Completed Treatment Episodes") +
  guides(fill = guide_legend(reverse = TRUE))
```