---
title: "Outcome Aggregation"
author: "Connor Flynn"
date: "11/28/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(haven)
library(feather)
library(arrow)
options(scipen = 99)

```

Load in Data

```{r}
teds_d <- arrow::read_feather(here("data/tedsD_2012_2020.feather"))
```


Clean Column Names

```{r}
teds_d <- teds_d %>% 
  clean_names()
```


Filter for years with admission/discharge data


```{r}
teds_d_select_years <- teds_d %>% 
  filter(disyr %in% c("2015", "2016", "2017", "2018", "2019", "2020"))
```




Replace -9 with NA

```{r}
teds_d_select_years[teds_d_select_years == "-9"] <- NA
```




Filter for completed treatment

```{r}
teds_d_select_years_completed <- teds_d_select_years %>% 
  filter(reason == 1)
```


Holistic Improvements 

Join admission and discharge columns


Add Employment Change 

```{r}
teds_d_select_years_completed <- teds_d_select_years_completed %>% 
  mutate(employment_change_specific = paste(employ, employ_d, sep="-"))
```


```{r}
teds_d_select_years_completed %>% 
  select(employ, employ_d, employment_change_specific)
```

Add Frequency of Use Change

```{r}
teds_d_select_years_completed <- teds_d_select_years_completed %>% 
  mutate(frequency_of_use_change_specific = paste(freq1, freq1_d, sep="-"))
```


```{r}
teds_d_select_years_completed %>% 
  select(freq1, freq1_d, frequency_of_use_change_specific)
```


Add Living Arrangement Change

```{r}
teds_d_select_years_completed <- teds_d_select_years_completed %>% 
  mutate(living_arrangement_change_specific = paste(livarag, livarag_d, sep="-"))
```

```{r}
teds_d_select_years_completed %>% 
  select(livarag, livarag_d, living_arrangement_change_specific)
```


Add Self Help Change

```{r}
teds_d_select_years_completed <- teds_d_select_years_completed %>% 
  mutate(self_help_change_specific = paste(freq_atnd_self_help, freq_atnd_self_help_d, sep="-"))
```

```{r}
teds_d_select_years_completed %>% 
  select(freq_atnd_self_help, freq_atnd_self_help_d, self_help_change_specific)
```


Aggregate Outcomes

Frequency of Use Change

```{r}
teds_d_select_years_completed <- teds_d_select_years_completed %>%
  mutate(frequency_of_use_change_aggregate = case_when(
    frequency_of_use_change_specific %in% c("3-2", "2-1", "3-1") ~ "improvement",
    frequency_of_use_change_specific %in% c("3-3", "2-2") ~ "stagnant negative",
    frequency_of_use_change_specific == "1-1" ~ "stagnant positive",
    frequency_of_use_change_specific %in% c("1-2", "2-3", "1-3") ~ "regression",
    TRUE ~ "unknown"
  ))
```

```{r}
table(teds_d_select_years_completed$frequency_of_use_change_aggregate)
```



Employment Change

```{r}
teds_d_select_years_completed <- teds_d_select_years_completed %>%
  mutate(employment_change_aggregate = case_when(
    employment_change_specific %in% c("4-3", "4-2", "4-1", "3-2", "3-1", "2-1") ~ "improvement",
    employment_change_specific %in% c("4-4", "3-3") ~ "stagnant negative",
    employment_change_specific %in% c("1-1", "2-2") ~ "stagnant positive",
    employment_change_specific %in% c("1-2", "2-3", "3-4", "1-4", "1-3", "2-4") ~ "regression",
    TRUE ~ "unknown"
  ))
```

```{r}
table(teds_d_select_years_completed$employment_change_aggregate)
```

Living Arrangements


```{r}
teds_d_select_years_completed <- teds_d_select_years_completed %>%
  mutate(living_arrangement_change_aggregate = case_when(
    living_arrangement_change_specific %in% c("1-2", "1-3", "2-3") ~ "improvement",
    living_arrangement_change_specific == "1-1" ~ "stagnant negative",
    living_arrangement_change_specific == "3-3" ~ "stagnant positive",
    living_arrangement_change_specific == "2-2" ~ "neutral",
    living_arrangement_change_specific %in% c("2-1", "3-1", "3-2") ~ "regression",
    TRUE ~ "unknown"
  ))
```

```{r}
table(teds_d_select_years_completed$living_arrangement_change_aggregate)
```

Self Help

```{r}
teds_d_select_years_completed <- teds_d_select_years_completed %>%
  mutate(self_help_change_aggregate = case_when(
    self_help_change_specific %in% c("1-2", "2-3", "3-4", "1-5", "1-3", "1-4", "2-4") ~ "improvement",
    self_help_change_specific == "1-1" ~ "stagnant negative",
    self_help_change_specific %in% c("2-2", "3-3", "4-4", "5-5") ~ "stagnant positive",
    self_help_change_specific %in% c("4-3", "3-2", "2-1", "4-1", "4-2", "3-1" ) ~ "regression",
    TRUE ~ "unknown"
  ))
```

```{r}
teds_d_select_years_completed %>% 
  select(employ, employ_d, employment_change_specific, employment_change_aggregate,
         freq1, freq1_d, frequency_of_use_change_specific, frequency_of_use_change_aggregate,
         freq_atnd_self_help, freq_atnd_self_help_d, self_help_change_specific, self_help_change_aggregate,
         livarag, livarag_d, living_arrangement_change_specific, living_arrangement_change_aggregate)
```




Write Dataframe

```{r}
write_csv(teds_d_select_years_completed, here("data/teds_d_select_years_completed_outcomes_aggregated.csv"))

#teds_d_select_years_completed <- read_csv(here("teds_d_select_years_completed_outcomes_aggregated.csv"))
```