---
title: "partial dependence"
author: "Connor Flynn"
date: "1/9/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```


*State* Feature 4 (dropped from faceted plot due to large y axis for states)

```{r}
pdp_stfips <- read_csv(here("data/pdp_stfips.csv"))
```

Replace fips code with state name

```{r}
state_info <- data.frame(
  stfips = c(1, 2, 4, 5, 6, 8, 9, 10, 11, 12, 13, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 42, 44, 45, 46, 47, 48, 49, 50, 51, 55, 56, 72),
  State = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Wisconsin", "Wyoming", "Puerto Rico")
)
```

Merge Dataframes

```{r}
pdp_stfips_names <- merge(pdp_stfips, state_info, by = "stfips", all.x = TRUE)
```


Plot

```{r}
ggplot(pdp_stfips_names, aes(y = reorder(State, yhat), x = yhat)) +
  geom_bar(stat = "identity", fill = "#639CA4FF", color = "black") +
  theme_minimal() +
  labs(title = "Partial Dependence Plot",
       x = "Prediction of No-Use at Discharge",
       y = "")
```



*Freq1* Feature 1
Frequency of Use at Admission

```{r}
pdp_freq1 <- read_csv(here("data/pdp_freq1.csv"))
```

```{r}
pdp_freq1$freq1[pdp_freq1$freq1 == "1"] <- "no use"
pdp_freq1$freq1[pdp_freq1$freq1 == "2"] <- "some use"
pdp_freq1$freq1[pdp_freq1$freq1 == "3"] <- "daily use"
```

```{r}
ggplot(pdp_freq1, aes(y = reorder(freq1, yhat), x = yhat)) +
  geom_bar(stat = "identity", fill = "#639CA4FF", color = "black") +
  theme_minimal() +
  labs(title = "Partial Dependence Plot",
       x = "Prediction of No-Use at Discharge",
       y = "")
```


*Length of Stay (Binned)* Feature 2


```{r}
pdp_los_binned <- read_csv(here("data/pdp_los_binned.csv"))
```


Plot

```{r}
ggplot(pdp_los_binned, aes(y = reorder(los_binned, yhat), x = yhat)) +
  geom_bar(stat = "identity", fill = "#639CA4FF", color = "black") +
  theme_minimal() +
  labs(title = "Partial Dependence Plot",
       x = "Prediction of No-Use at Discharge",
       y = "")
```




*Services at Discharge* Feature 3

```{r}
pdp_services <- read_csv(here("data/pdp_services.csv"))
```

Replace services codes with label

```{r}
pdp_services$services_d[pdp_services$services_d == "1"] <- "Detox, 24-hour, hospital inpatient"

pdp_services$services_d[pdp_services$services_d == "2"] <- "Detox, 24-hour, free-standing residential"

pdp_services$services_d[pdp_services$services_d == "3"] <- "Rehab/residential, hospital (non-detox)"

pdp_services$services_d[pdp_services$services_d == "4"] <- "Rehab/residential, short term (30 days or fewer)"

pdp_services$services_d[pdp_services$services_d == "5"] <- "Rehab/residential, long term (more than 30 days)"

pdp_services$services_d[pdp_services$services_d == "6"] <- "Ambulatory, intensive outpatient"

pdp_services$services_d[pdp_services$services_d == "7"] <- "Ambulatory, non-intensive outpatient"

pdp_services$services_d[pdp_services$services_d == "8"] <- "Ambulatory, detoxification"

```




Plot

```{r}
ggplot(pdp_services, aes(y = reorder(services_d, yhat), x = yhat)) +
  geom_bar(stat = "identity", fill = "#639CA4FF", color = "black") +
  theme_minimal() +
  labs(title = "Partial Dependence Plot",
       x = "Prediction of No-Use at Discharge",
       y = "")
```

*Division* Feature 5

```{r}
pdp_division <- read_csv(here("data/pdp_division.csv"))
```

```{r}
pdp_division$division[pdp_division$division == "0"] <- "U.S. territories"

pdp_division$division[pdp_division$division == "1"] <- "New England"

pdp_division$division[pdp_division$division == "2"] <- "Middle Atlantic"

pdp_division$division[pdp_division$division == "3"] <- "East North Central"

pdp_division$division[pdp_division$division == "4"] <- "West North Central"

pdp_division$division[pdp_division$division == "5"] <- "South Atlantic"

pdp_division$division[pdp_division$division == "6"] <- "East South Central"

pdp_division$division[pdp_division$division == "7"] <- "West South Central"

pdp_division$division[pdp_division$division == "8"] <- "Mountain"

pdp_division$division[pdp_division$division == "9"] <- "Pacific"
```


```{r}
ggplot(pdp_division, aes(y = reorder(division, yhat), x = yhat)) +
  geom_bar(stat = "identity", fill = "#639CA4FF", color = "black") +
  theme_minimal() +
  labs(title = "Partial Dependence Plot",
       x = "Prediction of No-Use at Discharge",
       y = "")
```

*Region*

```{r}
pdp_region <- read_csv(here("data/pdp_region.csv"))
```

Replace services codes with label

```{r}
pdp_region$region[pdp_region$region == "0"] <- "U.S. territories"

pdp_region$region[pdp_region$region == "1"] <- "Northeast"

pdp_region$region[pdp_region$region == "2"] <- "Midwest"

pdp_region$region[pdp_region$region == "3"] <- "South"

pdp_region$region[pdp_region$region == "4"] <- "West"


```


Plot

```{r}
ggplot(pdp_region, aes(y = reorder(region, yhat), x = yhat)) +
  geom_bar(stat = "identity", fill = "#639CA4FF", color = "black") +
  theme_minimal() +
  labs(title = "Partial Dependence Plot",
       x = "Prediction of No-Use at Discharge",
       y = "")
```




*Living Arrangement Change Aggregate*

```{r}
pdp_living_arrangement_change_aggregate <- read_csv(here("data/pdp_living_arrangement_change_aggregate.csv"))
```


Plot

```{r}
ggplot(pdp_living_arrangement_change_aggregate, aes(y = reorder(living_arrangement_change_aggregate, yhat), x = yhat)) +
  geom_bar(stat = "identity", fill = "#639CA4FF", color = "black") +
  theme_minimal() +
  labs(title = "Partial Dependence Plot",
       x = "Prediction of No-Use at Discharge",
       y = "")
```


*Self Help at Discharge*

```{r}
pdp_self_help <- read_csv(here("data/pdp_self_help.csv"))
```


Replace services codes with label

```{r}
pdp_self_help$freq_atnd_self_help_d[pdp_self_help$freq_atnd_self_help_d == "1"] <- "No attendance"

pdp_self_help$freq_atnd_self_help_d[pdp_self_help$freq_atnd_self_help_d == "2"] <- "1-3 times in the past month"

pdp_self_help$freq_atnd_self_help_d[pdp_self_help$freq_atnd_self_help_d == "3"] <- "4-7 times in the past month"

pdp_self_help$freq_atnd_self_help_d[pdp_self_help$freq_atnd_self_help_d == "4"] <- "8-30 times in the past month"

pdp_self_help$freq_atnd_self_help_d[pdp_self_help$freq_atnd_self_help_d == "5"] <- "Some attendance, frequency is unknown"

```




Plot

```{r}
ggplot(pdp_self_help, aes(y = reorder(freq_atnd_self_help_d, yhat), x = yhat)) +
  geom_bar(stat = "identity", fill = "#639CA4FF", color = "black") +
  theme_minimal() +
  labs(title = "Partial Dependence Plot",
       x = "Prediction of No-Use at Discharge",
       y = "")
```




*DSMCRIT* heavily imputed (45%)

```{r}
pdp_dsmcrit <- read_csv(here("data/pdp_dsmcrit.csv"))
```



```{r}
diagnosis_mapping <- c(
  "1" = "Alcohol-induced disorder",
  "2" = "Substance-induced disorder",
  "3" = "Alcohol intoxication",
  "4" = "Alcohol dependence",
  "5" = "Opioid dependence",
  "6" = "Cocaine dependence",
  "7" = "Cannabis dependence",
  "8" = "Other substance dependence",
  "9" = "Alcohol abuse",
  "10" = "Cannabis abuse",
  "11" = "Other substance abuse",
  "12" = "Opioid abuse",
  "13" = "Cocaine abuse",
  "14" = "Anxiety disorders",
  "15" = "Depressive disorders",
  "16" = "Schizophrenia/other psychotic disorders",
  "17" = "Bipolar disorders",
  "18" = "Attention deficit/disruptive behavior disorders",
  "19" = "Other mental health condition",
  "-9" = "Missing/unknown/not collected/invalid/no or deferred diagnosis"
)

# Use the named vector to replace values with labels
pdp_dsmcrit$dsmcrit <- diagnosis_mapping[as.character(pdp_dsmcrit$dsmcrit)]

```


```{r}
ggplot(pdp_dsmcrit, aes(y = reorder(dsmcrit, yhat), x = yhat)) +
  geom_bar(stat = "identity", fill = "#639CA4FF", color = "black") +
  theme_minimal() +
  labs(title = "Partial Dependence Plot",
       x = "Prediction of No-Use at Discharge",
       y = "")
```

